---
title: "Two-way Fixed Effects and Event Studies"
subtitle: "<html><div style='float:left'></div><hr color='#EB811B' size=1px width=0px></html>"
author: Ian McCarthy | Emory University
date: Workshop on Causal Inference with Panel Data
header-includes: 
  - \usepackage{graphicx}
  - \usepackage{amsmath}
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, custom.css, cols.css] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: [macros.js, cols_macro.js]
---

<!-- Adjust some CSS code for font size and maintain R code font size -->
<style type="text/css">
.remark-slide-content {
    font-size: 30px;
    padding: 1em 2em 1em 2em;    
}
.remark-code {
  font-size: 15px;
}
.remark-inline-code { 
    font-size: 20px;
}
</style>


<!-- Set R options for how code chunks are displayed and load packages -->
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(knitr)
knitr::opts_chunk$set(
  fig.align="center",  
  fig.height=3, #fig.width=6,
  # out.width="748px", #out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T,# echo=F, warning=F, message=F
  warning = FALSE, 
  message = FALSE, 
  cache.lazy = FALSE,
  error=TRUE
  )

knitr::opts_hooks$set(fig.callout = function(options) {
  if(options$fig.callout) {
    options$echo = FALSE
  }
  options
})

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes,
               scales, plotly, gganimate, cobalt, stargazer, haven, ggthemes,
               gifski, magick, lfe, dotwhisker, here, modelsummary, fixest)
set.seed(1234)
```

# Table of contents

1. [Two-way Fixed Effects](#twfe)
2. [Event Studies](#event)
3. [In Practice](#handson)

---
class: inverse, center, middle
name: twfe

# The Idea of TWFE

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>

---
# What is TWFE?

- Just a shorthand for a common regression specification
- Fixed effects for each unit and each time period, $\lambda_{i}$ and $\lambda_{t}$
- More general than 2x2 DD but same result

---
count: false

# What is TWFE?
Want to estimate $\delta$:

$$y_{it} = \alpha + \delta D_{it} + \gamma_{i} + \gamma_{t} + \varepsilon,$$<br>

where $\gamma_{i}$ and $\gamma_{t}$ denote a set of unit $i$ and time period $t$ dummy variables (or fixed effects).

---
# TWFE in Practice
```{r}
library(tidyverse)
library(modelsummary)
mcaid.data <- read_tsv("https://raw.githubusercontent.com/imccart/empirical-methods/main/data/medicaid-expansion/mcaid-expand-data.txt")
reg.dat <- mcaid.data %>% filter(expand_year==2014 | is.na(expand_year), !is.na(expand_ever)) %>%
  mutate(perc_unins=uninsured/adult_pop,
         post = (year>=2014), 
         treat=post*expand_ever)
```

---
count: false

# TWFE in Practice

```{r}
summary(lm(perc_unins ~ post + expand_ever + post*expand_ever, data=reg.dat))
```

---
count: false

# TWFE in Practice

```{r}
library(fixest)
twfe <- feols(perc_unins ~ expand | State + year, data=reg.dat)
twfe$coeftable
```


---
class: inverse, center, middle
name: event

# Event Studies

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>

---
# What is an event study?

Event study is poorly named:
- In finance, even study is just an *interrupted time series*
- In econ and other areas, we usually have a treatment/control group *and* a break in time

---
count: false

# What is an event study?

- Allows for different effect estimates at each time period (maybe effects phase in over time or dissipate)
- Visually very appealing
- Offers easy "smell test" for parallel trends assumption


---
count: false

# What is an event study?

Estimate something akin to...
$$y_{it} = \gamma_{i} + \gamma_{t} + \sum_{\tau = -q}^{-1}\delta_{\tau} D_{i \tau} + \sum_{\tau=0}^{m} \delta_{\tau}D_{i \tau} + x_{it} + \epsilon_{it},$$

where $q$ captures the number of periods before the treatment occurs and $m$ captures periods after treatment occurs.

---
# How to do an event study?

First create all of the treatment/year interactions:
```{r}
event.dat <- reg.dat %>%
  mutate(expand_2012 = expand_ever*(year==2012),
         expand_2013 = expand_ever*(year==2013),
         expand_2014 = expand_ever*(year==2014),
         expand_2015 = expand_ever*(year==2015),
         expand_2016 = expand_ever*(year==2016),
         expand_2017 = expand_ever*(year==2017),
         expand_2018 = expand_ever*(year==2018))
```

---
count: false

# How to do an event study?

Second, run regression with full set of interactions and group/year dummies:
```{r}
event.ins.reg <- lm(perc_unins ~ expand_2012 + expand_2014 + 
                      expand_2015 + expand_2016 + expand_2017 + 
                      expand_2018 + factor(year) + factor(State), data=event.dat)
point.est <- as_tibble(c(event.ins.reg$coefficients[c("expand_2012","expand_2014","expand_2015",
                                            "expand_2016","expand_2017","expand_2018")]),
                       rownames = "term")
ci.est <- as_tibble(confint(event.ins.reg)[c("expand_2012","expand_2014","expand_2015",
                                   "expand_2016","expand_2017","expand_2018"),],
                    rownames = "term")
```

---
count: false

# How to do an event study?

Third, organize results into a new dataset:
```{r}
point.est <- point.est %>% rename(estimate = value)
ci.est <- ci.est %>% rename(conf.low = `2.5 %`, conf.high = `97.5 %`)
new.row <- tibble(
  term = "expand_2013",
  estimate = 0,
  conf.low = 0,
  conf.high = 0,
  year = 2013
)

event.plot.dat <- point.est %>%
  left_join(ci.est, by=c("term")) %>%
  mutate(year = c(2012, 2014, 2015, 2016, 2017, 2018)) %>%
  bind_rows(new.row) %>%
  arrange(year)
```

---
count: false

# How to do an event study?

Finally, plot coefficients and confidence intervals
```{r es-plot, warning=FALSE, include=FALSE}
dwplot(event.plot.dat, 
       vline=geom_vline(xintercept=0, linetype=2), 
       order_vars = c("expand_2012","expand_2013","expand_2014",
                      "expand_2015","expand_2016","expand_2017",
                      "expand_2018"),
       whisker_args = list(color="black", size=1.1),
       dot_args = list(color="black")) + 
  coord_flip() + theme_bw() + theme(legend.position = "none") +
  labs(y = "Year",
       x = "Estimate and 95% CI",
       title = "Event Study Estimates for Medicaid and Uninsurance Rate") +
  scale_y_discrete(labels = c("expand_2018" = "2018", 
                              "expand_2017" = "2017",
                              "expand_2016" = "2016",
                              "expand_2015" = "2015",
                              "expand_2014" = "2014",
                              "expand_2013" = "2013",
                              "expand_2012" = "2012"))
```

.plot-callout[
```{r es-plot-small, ref.label="es-plot", fig.callout=TRUE, warning=FALSE}
```
]

---
count: false

# How to do an event study?

```{r es-plot-big, ref.label="es-plot", fig.callout=TRUE, warning=FALSE}
```


---
class: inverse, center, middle
name: handson

# Seeing things in action

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>
