---
title: "Two-way Fixed Effects and Event Studies"
subtitle: "<html><div style='float:left'></div><hr color='#EB811B' size=1px width=0px></html>"
author: Ian McCarthy | Emory University
date: Workshop on Causal Inference with Panel Data
header-includes: 
  - \usepackage{graphicx}
  - \usepackage{amsmath}
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, custom.css, cols.css] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      beforeInit: [macros.js, cols_macro.js]
---

<!-- Adjust some CSS code for font size and maintain R code font size -->
<style type="text/css">
.remark-slide-content {
    font-size: 30px;
    padding: 1em 2em 1em 2em;    
}
.remark-code {
  font-size: 15px;
}
.remark-inline-code { 
    font-size: 20px;
}
</style>


<!-- Set R options for how code chunks are displayed and load packages -->
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(knitr)
knitr::opts_chunk$set(
  fig.align="center",  
  fig.height=3, #fig.width=6,
  # out.width="748px", #out.length="520.75px",
  dpi=300, #fig.path='Figs/',
  cache=T,# echo=F, warning=F, message=F
  warning = FALSE, 
  message = FALSE, 
  cache.lazy = FALSE,
  error=TRUE
  )

knitr::opts_hooks$set(fig.callout = function(options) {
  if(options$fig.callout) {
    options$echo = FALSE
  }
  options
})

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, here, modelsummary, fixest, data.table,
               ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes,
               scales, plotly, gganimate, cobalt, ivpack, stargazer, haven, ggthemes,
               gifski, magick, lfe, dotwhisker)
set.seed(1234)
```

# Table of contents

1. [Two-way Fixed Effects](#twfe)
2. [Event Studies](#event)
3. [In Practice](#handson)
4. [What are we estimating?](#what)

---
class: inverse, center, middle
name: twfe

# The Idea of TWFE

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>

---
# What is TWFE?

- Just a shorthand for a common regression specification
- Fixed effects for each unit and each time period, $\lambda_{i}$ and $\lambda_{t}$
- More general than 2x2 DD but same result

---
count: false

# What is TWFE?
Want to estimate $\delta$:

$$y_{it} = \alpha + \delta D_{it} + \gamma_{i} + \gamma_{t} + \varepsilon,$$<br>

where $\gamma_{i}$ and $\gamma_{t}$ denote a set of unit $i$ and time period $t$ dummy variables (or fixed effects).

---
# TWFE in Practice
```{r}
library(tidyverse)
library(modelsummary)
mcaid.data <- read_tsv("https://raw.githubusercontent.com/imccart/empirical-methods/main/data/medicaid-expansion/mcaid-expand-data.txt")
reg.dat <- mcaid.data %>% filter(expand_year==2014 | is.na(expand_year), !is.na(expand_ever)) %>%
  mutate(perc_unins=uninsured/adult_pop,
         post = (year>=2014), 
         treat=post*expand_ever)
```

---
count: false

# TWFE in Practice

```{r}
summary(lm(perc_unins ~ post + expand_ever + post*expand_ever, data=reg.dat))
```

---
count: false

# TWFE in Practice

```{r}
library(fixest)
twfe <- feols(perc_unins ~ expand | State + year, data=reg.dat)
twfe$coeftable
```


---
class: inverse, center, middle
name: event

# Event Studies

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>

---
# What is an event study?

Event study is poorly named:
- In finance, even study is just an *interrupted time series*
- In econ and other areas, we usually have a treatment/control group *and* a break in time

---
count: false

# What is an event study?

- Allows for different effect estimates at each time period (maybe effects phase in over time or dissipate)
- Visually very appealing
- Offers easy "smell test" for parallel trends assumption


---
count: false

# What is an event study?

Estimate something akin to...
$$y_{it} = \gamma_{i} + \gamma_{t} + \sum_{\tau = -q}^{-1}\delta_{\tau} D_{i \tau} + \sum_{\tau=0}^{m} \delta_{\tau}D_{i \tau} + x_{it} + \epsilon_{it},$$

where $q$ captures the number of periods before the treatment occurs and $m$ captures periods after treatment occurs.

---
# How to do an event study?

First create all of the treatment/year interactions:
```{r}
event.dat <- reg.dat %>%
  mutate(expand_2012 = expand_ever*(year==2012),
         expand_2013 = expand_ever*(year==2013),
         expand_2014 = expand_ever*(year==2014),
         expand_2015 = expand_ever*(year==2015),
         expand_2016 = expand_ever*(year==2016),
         expand_2017 = expand_ever*(year==2017),
         expand_2018 = expand_ever*(year==2018))
```

---
count: false

# How to do an event study?

Second, run regression with full set of interactions and group/year dummies:
```{r}
event.ins.reg <- lm(perc_unins ~ expand_2012 + expand_2014 + 
                      expand_2015 + expand_2016 + expand_2017 + 
                      expand_2018 + factor(year) + factor(State), data=event.dat)
point.est <- as_tibble(c(event.ins.reg$coefficients[c("expand_2012","expand_2014","expand_2015",
                                            "expand_2016","expand_2017","expand_2018")]),
                       rownames = "term")
ci.est <- as_tibble(confint(event.ins.reg)[c("expand_2012","expand_2014","expand_2015",
                                   "expand_2016","expand_2017","expand_2018"),],
                    rownames = "term")
```

---
count: false

# How to do an event study?

Third, organize results into a new dataset:
```{r}
point.est <- point.est %>% rename(estimate = value)
ci.est <- ci.est %>% rename(conf.low = `2.5 %`, conf.high = `97.5 %`)
new.row <- tibble(
  term = "expand_2013",
  estimate = 0,
  conf.low = 0,
  conf.high = 0,
  year = 2013
)

event.plot.dat <- point.est %>%
  left_join(ci.est, by=c("term")) %>%
  mutate(year = c(2012, 2014, 2015, 2016, 2017, 2018)) %>%
  bind_rows(new.row) %>%
  arrange(year)
```

---
count: false

# How to do an event study?

Finally, plot coefficients and confidence intervals
```{r es-plot, warning=FALSE, include=FALSE}
dwplot(event.plot.dat,
       vline=geom_vline(xintercept=0, linetype=2), 
       vars_order = c("expand_2018","expand_2017","expand_2016",
                      "expand_2015","expand_2014","expand_2013",
                      "expand_2012"),
       whisker_args = list(color="black", size=1.1),
       dot_args = list(color="black")) + 
  coord_flip() + theme_bw() + theme(legend.position = "none") +
  labs(y = "Year",
       x = "Estimate and 95% CI",
       title = "Event Study Estimates for Medicaid and Uninsurance Rate") +
  scale_y_discrete(labels = c("expand_2018" = "2018", 
                              "expand_2017" = "2017",
                              "expand_2016" = "2016",
                              "expand_2015" = "2015",
                              "expand_2014" = "2014",
                              "expand_2013" = "2013",
                              "expand_2012" = "2012"))
```

.plot-callout[
```{r es-plot-small, ref.label="es-plot", fig.callout=TRUE, warning=FALSE}
```
]

---
count: false

# How to do an event study?

```{r es-plot-big, ref.label="es-plot", fig.callout=TRUE, warning=FALSE}
```


---
class: inverse, center, middle
name: handson

# Seeing things in action

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>


---
# Things to address

1. "Event time" vs calendar time
2. Define baseline period
3. Choose number of pre-treatment and post-treatment coefficients


---
# Event time vs calendar time

Essentially two "flavors" of event studies

1. Common treatment timing
2. Differential treatment timing

---
# Define baseline period

- Must choose an "excluded" time period (as in all cases of group dummy variables)
- Common choice is $t=-1$ (period just before treatment)
- Easy to understand with calendar time
- For event time...manually set time to $t=-1$ for all untreated units

---
# Number of pre-treatment and post-treatment periods

- On event time, sometimes very few observations for large lead or lag values
- Medicaid expansion example: Late adopting states have fewer post-treatment periods
- Norm is to group final lead/lag periods together


---
# Common treatment timing

.pull-left[
**Stata**<br>
```{stata, eval=FALSE}
ssc install reghdfe

insheet using "https://raw.githubusercontent.com/imccart/empirical-methods/main/data/medicaid-expansion/mcaid-expand-data.txt", clear
gen perc_unins=uninsured/adult_pop
keep if expand_year=="2014" | expand_year=="NA"
drop if expand_ever=="NA"
gen post=(year>=2014)
gen treat=(expand_ever=="TRUE")
gen treat_post=(expand=="TRUE")

reghdfe perc_unins treat##ib2013.year, absorb(state)
gen coef = .
gen se = .
forvalues i = 2012(1)2018 {
    replace coef = _b[1.treat#`i'.year] if year == `i'
    replace se = _se[1.treat#`i'.year] if year == `i'
}

* Make confidence intervals
gen ci_top = coef+1.96*se
gen ci_bottom = coef - 1.96*se

* Limit ourselves to one observation per year
keep year coef se ci_*
duplicates drop

* Create connected scatterplot of coefficients
* with CIs included with rcap 
* and a line at 0 from function
twoway (sc coef year, connect(line)) (rcap ci_top ci_bottom year) ///
    (function y = 0, range(2012 2018)), xtitle("Year") ///
    caption("Estimates and 95% CI from Event Study")
```
]


.pull-right[
**R**<br>
```{r, eval=FALSE}
library(tidyverse)
library(modelsummary)
library(data.table)
library(fixest)
mcaid.data <- read_tsv("https://raw.githubusercontent.com/imccart/empirical-methods/main/data/medicaid-expansion/mcaid-expand-data.txt")
reg.dat <- as.data.table(mcaid.data) %>% 
  filter(expand_year==2014 | is.na(expand_year), !is.na(expand_ever)) %>%
  mutate(perc_unins=uninsured/adult_pop,
         post = (year>=2014), 
         treat=post*expand_ever)

mod.twfe <- feols(perc_unins~i(year, expand_ever, ref=2013) | State + year,
                  cluster=~State,
                  data=reg.dat)
iplot(mod.twfe, 
      xlab = 'Time to treatment',
      main = 'Event study')
```
]


---
# Comparing results

```{r es-plot-iplot, warning=FALSE, include=FALSE}
reg.dat <- as.data.table(mcaid.data) %>% 
  filter(expand_year==2014 | is.na(expand_year), !is.na(expand_ever)) %>%
  mutate(perc_unins=uninsured/adult_pop,
         post = (year>=2014), 
         treat=post*expand_ever)

mod.twfe <- feols(perc_unins~i(year, expand_ever, ref=2013) | State + year,
                  cluster=~State,
                  data=reg.dat)
iplot(mod.twfe, 
      xlab = 'Time to treatment',
      main = 'Event study',
      ylab = "Estimate and 95% CI")
```

```{r es-plot2, ref.label="es-plot-iplot", fig.callout=TRUE, warning=FALSE, out.width="80%", fig.align='left'}
```

.plot-callout[
```{r es-plot1, ref.label="es-plot", fig.callout=TRUE, warning=FALSE}
```
]


---
# Differential treatment timing

- Now let's work with the full Medicaid expansion data
- Includes late adopters
- Requires putting observations on "event time"

---
count: false

# Differential treatment timing

.pull-left[
**Stata**<br>
```{stata, eval=FALSE}
ssc install reghdfe

insheet using "https://raw.githubusercontent.com/imccart/empirical-methods/main/data/medicaid-expansion/mcaid-expand-data.txt", clear
gen perc_unins=uninsured/adult_pop
drop if expand_ever=="NA"
replace expand_year="." if expand_year=="NA"
destring expand_year, replace
gen event_time=year-expand_year
replace event_time=-1 if event_time==.

forvalues l = 0/4 {
	gen L`l'event = (event_time==`l')
}
forvalues l = 1/2 {
	gen F`l'event = (event_time==-`l')
}
gen F3event=(event_time<=-3)

reghdfe perc_unins F3event F2event L0event L1event L2event L3event L4event, absorb(state year) cluster(state)
gen coef = .
gen se = .
forvalues i = 2(1)3 {
    replace coef = _b[F`i'event] if F`i'event==1
    replace se = _se[F`i'event] if F`i'event==1
}
forvalues i = 0(1)4 {
    replace coef = _b[L`i'event] if L`i'event==1
    replace se = _se[L`i'event] if L`i'event==1
}
replace coef = 0 if F1event==1
replace se=0 if F1event==1

* Make confidence intervals
gen ci_top = coef+1.96*se
gen ci_bottom = coef - 1.96*se

* Limit ourselves to one observation per year
keep if event_time>=-3 & event_time<=4
keep event_time coef se ci_*
duplicates drop

* Create connected scatterplot of coefficients
* with CIs included with rcap 
* and a line at 0 from function
sort event_time
twoway (sc coef event_time, connect(line)) (rcap ci_top ci_bottom event_time) ///
    (function y = 0, range(-3 4)), xtitle("Time") ///
    caption("Estimates and 95% CI from Event Study") xlabel(-3(1)4)
```
]

.pull-right[
**R**<br>
```{r, eval=FALSE}
library(tidyverse)
library(modelsummary)
library(data.table)
library(fixest)
mcaid.data <- read_tsv("https://raw.githubusercontent.com/imccart/empirical-methods/main/data/medicaid-expansion/mcaid-expand-data.txt")
reg.dat <- as.data.table(mcaid.data) %>% 
  filter(!is.na(expand_ever)) %>%
  mutate(perc_unins=uninsured/adult_pop,
         post = (year>=2014), 
         treat=post*expand_ever)

reg.dat[, time_to_treat := ifelse(expand_ever==TRUE, year-expand_year, 0)]
mod.twfe <- feols(perc_unins~i(time_to_treat, expand_ever, ref=-1) | State + year,
                  cluster=~State,
                  data=reg.dat)
iplot(mod.twfe, 
      xlab = 'Time to treatment',
      main = 'Event study')
```
]

---
class: inverse, center, middle
name: what

# What are we estimating?

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=1055px></html>

---
# Problems with TWFE

- Recall goal of estimating ATE or ATT
- TWFE and 2x2 DD identical with homogeneoues effects and common treatment timing
- Otherwise...TWFE is biased and inconsistent for ATT

---
# Inutition

- OLS is a weighted average of all 2x2 DD groups
- Weights are function of size of subsamples, size of treatment/control units, and timing of treatment
- Units treated in middle of sample receive larger weights
- Prior-treated units act as controls for late-treated units


--
Just the length of the panel will change the estimate!

---
# Does it really matter?

- Definitely! But how much?
- Large treatment effects for early treated units could reverse the sign of final estimate
- Let's explore this nice Shiny app from Kyle Butts: [Bacon-Decomposition Shiny App](https://kyle-butts.shinyapps.io/did_bacon/).
